### 2020-04-29 review/reviews.tex
Added reviewer summaries

\documentclass{article}

\begin{document}

\section{Reviewer 1}

The reviewer seemed mostly content with the submission, suggesting some additional references, typo corrections and some specific suggestions for rephrasing some parts to improve the paper's accuracy. 
Furthermore, they suggest to reframe the paper in the micro-navigation context to make it more appealing to researchers who wish to know the accuracy of bone-conduction guidance cues.
Finally, they also suggest that the variance in the partially sighted group need to be better explained, by adding their demographics for example.

\section{Reviewer 2}

\section{Reviewer 3}

\end{document}